A molecular dynamics study of the solvation of carbon dioxide in the ionic liquids \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]}

### 2018-09-07 ionic_liquids.tex
Review of Force Fields section

The versatility of ILs for different applications is in part due to the fact that their physical properties strongly depend on the particular cation-anion combination.
On the other hand, this feature also explains why the desirable transferability of force fields is particularly difficult to attain.
Indeed, it is a common practice to extract the potential parameters of the intramolecular (bond stretching, angle bending, torsion rotation) and Lennard-Jones (LJ) interactions from a well established force field and, then, only derive the partial atomic charges with the aim of ``adjusting'' the model for a specific pair of ions.
This procedure is, however, inaccurate as in conventional force fields the determination of the atomic charges precedes the dihedrals parametrization.
Another ad-hoc strategy, which intends to crudely incorporate polarization effects, is to scale the partial charges uniformly with a factor which normally takes values in the range 0.8-0.9.
This has proven to be effective for predicting dynamic properties \cite{Schr_der_2012}.

In this paper, we consider four models of \ce{[emim][B(CN)_4]}.
Koller \textit{et al.} \cite{Koller_2012} were the first to introduce a specific non-polarizable force field for that IL.
They tested three variations of a \textit{united-atoms} model, in which the alkyl groups \ce{CH_3} and \ce{CH_2} of \ce{[emim]^+} are replaced by pseudo-atoms.
Among them, the ``FF-3'' model best reproduces the experimental values of density, self-diffusion coefficients and viscosity.
In short, both the intramolecular and LJ parameters of the cation were taken from the work of Liu \textit{et al.} \cite{Liu_2006}, which in turn is based on the AMBER \cite{Cornell_1995} framework.
The intramolecular parameters involving the boron atom were extracted from the DREIDING \cite{Mayo_1990} force field and the anionic vdW parameters were refined based on the values reported by Price \textit{et al.} \cite{Price_2001}.
The atomic partial charges were derived using the EA-RESP method \cite{Basma_2001} from \textit{Post-Hartree-Fock ab-initio} calculations at MP2/6-31G$^\ast$+ level of theory, which automatically resulted in total charges on the ions of $\pm$0.8426$e$.

The potential parameters of \ce{[B(CN)_4]^-} derived by Koller \textit{et al.} \cite{Koller_2012} were then used by Batista \textit{et al.} \cite{Batista_2015} to study aqueous solutions of \ce{[emim][B(CN)_4]}.
In this case, the authors considered the \textit{all-atom} model of \ce{[emim]^+} developed by Cadena and Maginn \cite{Cadena_2006}.
The atomic partial charges were obtained using the CHELP procedure \cite{Breneman_1990}, resulting in net charges of $\pm$0.889$e$ on the ions.
It should be noted that the factors for scaling the 1-4 non-bonded interactions are different for these force fields: Cadena and Maginn \cite{Cadena_2006} apply factors of 1.0 and 0.4 for the LJ and Coulombic interactions, respectively; while in Koller \textit{et al.} \cite{Koller_2012} the corresponding values are 0.5 and 5/6, in accordance with the AMBER force field.
Batista \textit{et al.} \cite{Batista_2015} used the first ones (J.P. Coutinho, personal communication).

Another set of parameters for \ce{[emim][B(CN)_4]} was derived by Liu \textit{et al.} \cite{Liu_2014} in a study that investigates the dynamics of \ce{CO_2} and \ce{N_2} in that IL.
Parameters corresponding to the intramolecular and LJ interactions were extracted from GAFF \cite{Wang_2004}, while the atomic partial charges were computed using the RESP method \cite{Bayly_1993} from \textit{DFT}/B3LYP/6-311+G(d) calculations.
The net charges of the ions were manually adjusted to a value of $\pm$0.8$e$.

We also consider the parametrization reported by Weber and Kirchner \cite{Weber_2016}, which is based on the force field developed by P{\'{a}}dua and Canongia Lopes~\cite{Canongia_Lopes_2006}.
The potential parameters for the anion are not specified in Refs.~\citenum{Weber_2016} and~\citenum{Canongia_Lopes_2006}, and we obtained them by a direct communication with the authors of Ref.~\citenum{Weber_2016}.
The atomic partial charges were derived using the RESP procedure from \textit{Hartree-Fock} calculations at 6-31++G$^{\ast \ast}$ level of theory, resulting in net charges of $\pm$0.8$e$.

Finally, for \ce{[emim][NTf_2]} we use the model of K\"{o}ddermann~\textit{et al.} \cite{K_ddermann_2007}, which was derived by refining the force field due to P{\'{a}}dua and Canongia Lopes \cite{Canongia_Lopes_2006}.
This refining was characterized by their authors as ``unusual'', given that the LJ parameters were modified, while keeping the partial charges at their original values.
This model was recently employed by Kerl \textit{et al.} \cite{Kerl__2017}, who were able to accurately predict Henry constants of \ce{CO_2} at several temperatures.

@@

The solutes considered in this work are hexane, benzene, \ce{CO_2}, ethanol, cyclohexane and water.
Hexane and benzene were modeled using the OPLS-AA force field \cite{Jorgensen_1996}, while the TRAPPE potential \cite{Chen_2001,Potoff_2001} was employed for ethanol and \ce{CO_2}.
Cyclohexane was simulated using a rigid-body model \cite{munoz2015lennard} whose sites only interact via a LJ potential
For water, we employ the TIP4P/2005 model \cite{Abascal_2005,Vega_2011}.

>>>>

The versatility of ILs for different applications is in part due to the fact that their physical properties strongly depend on the particular cation-anion combination.
On the other hand, this feature also explains why the desirable transferability of force fields is particularly difficult to attain.
Indeed, it is a common practice to extract the potential parameters of the intramolecular (bond stretching, angle bending, torsion rotation) and Lennard-Jones (LJ) interactions from a well established force field and, then, only derive the partial atomic charges with the aim of adjusting the model for a specific pair of ions.
This procedure is, however, inconsistent with conventional force field parameterization in which the determination of atomic charges precedes the establishment of torsional terms.
Another ad-hoc strategy, which intends to roughly incorporate polarization effects, is to scale the partial charges uniformly with a factor which normally takes values in the range from 0.8 to 0.9.
This has proven to be effective for predicting dynamic properties \cite{Schr_der_2012}.

In this paper, we analyze four different models for \ce{[emim][B(CN)_4]}.
Koller \textit{et al}. \cite{Koller_2012} were the first to introduce a specific non-polarizable force field for this IL.
They tested three variations of a united-atoms model, in which the hydrogen atoms of alkyl groups \ce{CH_3} and \ce{CH_2} of cation \ce{[emim]^+} are treated implicitly.
Among these variations, the one named FF-3 best reproduces the experimental values of density, self-diffusion coefficients, and viscosity.
In short, Koller \textit{et al}. \cite{Koller_2012} took both the intramolecular and LJ parameters for the cation from the work of Liu \textit{et al}. \cite{Liu_2006}, which in turn based their model on the AMBER \cite{Cornell_1995} framework.
The intramolecular parameters involving the boron atom were extracted from the DREIDING \cite{Mayo_1990} force field and the anionic van der Waals parameters were refined based on the values reported by Price \textit{et al.} \cite{Price_2001}.
The atomic partial charges were derived using the EA-RESP method \cite{Basma_2001} from Post-Hartree-Fock \textit{ab-initio} calculations at MP2/6-31G\textsuperscript{*}+ level of theory, which automatically resulted in total charges on the ions of $\pm0.8426~\mathrm{e}$.

Batista \textit{et al}. \cite{Batista_2015} then used the parameters of Ref.~\citenum{Koller_2012} for \ce{[B(CN)_4]^-} to study aqueous solutions of \ce{[emim][B(CN)_4]}.
In this case, the authors considered the all-atom model of Cadena and Maginn \cite{Cadena_2006} for \ce{[emim]^+}.
The atomic partial charges were obtained using the CHELP procedure \cite{Breneman_1990}, resulting in net charges of $\pm 0.889~\mathrm{e}$.
It should be noted that the factors for scaling the 1-4 non-bonded interactions are different for these force fields.
Cadena and Maginn \cite{Cadena_2006} apply factors of 1.0 and 0.4 for the LJ and Coulombic interactions, respectively, while for Koller \textit{et al}. \cite{Koller_2012} the corresponding values are 1/2 and 5/6, in accordance with the AMBER force field.
Batista \textit{et al}. \cite{Batista_2015} used the former throughout (J. Coutinho, personal communication).

Another set of parameters for \ce{[emim][B(CN)_4]} was derived by Liu \textit{et al}. \cite{Liu_2014} in a study that investigates the dynamics of \ce{CO_2} and \ce{N_2} dissolved in this ionic liquid.
Parameters corresponding to the intramolecular and LJ interactions were extracted from GAFF \cite{Wang_2004}, while the atomic partial charges were computed using the RESP method \cite{Bayly_1993} with B3LYP/6-311+G(d) DFT calculations.
The net charges of the ions were manually adjusted to $\pm 0.8 \mathrm{e}$.

Finally, we also consider the parameterization which Weber and Kirchner \cite{Weber_2016} carried out for \ce{[emim][B(CN)_4]} based on the force field developed by P{\'{a}}dua and Canongia Lopes \cite{Canongia_Lopes_2006}.
As the potential parameters for the anion were specified neither in Ref.~\citenum{Weber_2016} nor in Ref.~\citenum{Canongia_Lopes_2006}, we obtained them by direct communication with the authors of Ref.~\citenum{Weber_2016}.
The atomic partial charges were derived using the RESP procedure with Hartree-Fock calculations at the \mbox{6-31++G\textsuperscript{**}} level of theory, resulting in net charges of $\pm 0.8 \mathrm{e}$.

In the case of ionic liquid \ce{[emim][NTf_2]}, we use the model of K\"{o}ddermann \textit{et al}. \cite{K_ddermann_2007}, which was also derived by refining parameters of the force field by P{\'{a}}dua and Canongia Lopes \cite{Canongia_Lopes_2006}.
The authors characterized this refinement as unusual, given that they modified the LJ parameters while keeping the partial charges unchanged.
Kerl \textit{et al}. \cite{Kerl__2017} recently employed this model to accurately predict Henry constants of \ce{CO_2} in \ce{[emim][NTf_2]} at several temperatures.

@@

Besides \ce{CO_2}, the solutes considered in this work are hexane, benzene, ethanol, cyclohexane, and water.
Hexane and benzene were modeled using the OPLS-AA force field \cite{Jorgensen_1996}, while the TRAPPE potential \cite{Chen_2001,Potoff_2001} was employed for ethanol and \ce{CO_2}.
Cyclohexane was simulated using a rigid-body model \cite{munoz2015lennard} whose sites interact via a LJ potential only.
For water, we employ the TIP4P/2005 model \cite{Abascal_2005,Vega_2011}.
